---
title: "Anomaly detection in R"
author: "Edwin Mutuma"
date: "7/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Importing And Previewing The Data                 
### Importing the data

```{r}
library(tidyverse)
```

```{r}
library(anomalize)
```

```{r}
library(tibbletime)
```

```{r}
library(dplyr)
```

```{r}
library(magrittr)
```

```{r}
path <- 'http://bit.ly/CarreFourSalesDataset' 
```

          
```{r}
# Grouping by server and converting to tibbletime

sales <- read.csv(path) %>% group_by(Date) %>%
   mutate(Date = as.Date (Date, format="%m/%d/%Y")) %>% as_tbl_time(index = Date)

head(sales)

```

```{r}
class(sales)
```

```{r}
str(sales)
```

```{r}
colSums(is.na(sales))
```

```{r}
# performing anomaly detection on the sales 
sales_anomalized <- sales %>% summarise(Orders = sum(Sales)) %>%
    time_decompose(Orders) %>%
    anomalize(remainder, alpha = 0.25, max_anoms = 0.2) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE) +
    ggtitle("alpha = 0.25, max_anoms = 0.2")
    
sales_anomalized
```

The default Alpha of 0.05 couldn't detect any anomalies but upon increasing the alpha to 0.25 I discovered some potential anomalies.

## **CONCLUSION**
* Most anomalies occurred in the months of February and March.
* Least anomalies occurred in the month of January.

## **RECOMMENDATIONS**
* The management should take a closer look into the sales that occur in the months of February and March as there is potential for either losses or pilfering of unexplainable profits.
* There should also be a system in place that detects the anomalies as soon as they occur so that they can be investigated as soon as they occur.