---
title: "Association Rules in R"
author: "Edwin Mutuma"
date: "7/16/2021"
output: html_document
---


## Importing And Previewing The Data          
### Importing the data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(arules)
```


## READING DATA
```{r}
df2 <- read.transactions("http://bit.ly/SupermarketDatasetII", sep = ",")
```

### Checking the data
```{r}
head(df2)
```

```{r}
tail(df2)
```

```{r}
str(df2)
```

```{r}
dim(df2)
```

```{r}
summary(df2)
```


### Cleaning the data
```{r}
# Previewing our first 5 transactions
#
inspect(df2[1:5])
```


## IMPLEMENTING THE SOLUTION
### Association Rules
```{r}
# To preview the items that make up our dataset,
# 
items<-as.data.frame(itemLabels(df2))
colnames(items) <- "Item"
head(items, 10)  
```

```{r}
# Generating a summary of the dataset
summary(df2)
```

```{r}
# Exploring the frequency of some articles 
itemFrequency(df2[, 8:10],type = "absolute")
round(itemFrequency(df2[, 8:10],type = "relative")*100,2)
```

```{r}
# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%
# 
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(df2, topN = 10,col="darkgreen")
itemFrequencyPlot(df2, support = 0.1,col="darkred")
```

```{r}
# Building a model based on association rules 
# using the apriori algorithm
# We use Min Support as 0.001 and confidence as 0.8
# ---
# 
rules <- apriori (df2, parameter = list(supp = 0.001, conf = 0.8))
rules
```

```{r}
# We can perform an exploration of our model 
# through the use of the summary function
summary(rules)
```

```{r}
# Ordering these rules by level of confidence
# then looking at the first ten rules.
# ---
# 
rules<-sort(rules, by="lift", decreasing=TRUE)
inspect(rules[1:10])
```

```{r}
# Ordering these rules by level of confidence
# then looking at the first ten rules.
# ---
# 
rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:10])
```

## *RECOMMENDATIONS*
We fail to use confidence as a way of measuring the association since it misrepresents the importance of an association. That's why we use lift as a way of measuring the strength of the rules.

1. Most rules had 4 items in each transaction, followed by rules with 5 and 3 items.

2. The top 5 items purchased include: mineral water, eggs, pasta, french fries and chocolate.

3. Transactions with 1 item only were 1754, and were the highest in number followed by transactions with 2 and 3 items.

4. Going by confidence, cake, eggs, pasta, meatballs, mineral water and milk should be placed in aisles that are not far away from each other since there is a pretty high confidence that people will tend to buy those items together.

5. Foods rich in proteins should also be placed closely together since customers tend to buy most of them together.

6. Escalope should have a targeted promotion especially with customers that purchased french fries,mushroom cream sauce and pasta.
